---
title: "Special Work Performance 2: Describing Data"
geometry: margin=2cm
output: 
  pdf_document:
    fig_caption: true
sansfont: times
fontsize: 12pt
---
\begin{center}
Humboldt University Berlin 

Customer Analytics and Customer Insights WS 2017/18 
by Prof. Dr. Daniel Klapper 

\textbf{Group 9:}\\  
\textbf{Natalie Habib}, 587954; 
\textbf{Julia Dullin}, 587779; \\
Submitted on November 28, 2017 

\end{center}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

salaries = read.csv("salaries.csv")
salaries.tmp <- sapply(salaries,is.numeric)
for(clm in salaries.tmp) 
{
  salaries$clm <- factor(clm)
}
salaries.df <- data.frame(salaries)

library("data.table")
library("gplots")
library("corrplot")
library(knitr)
library(car)
library(xtable)
library(lattice)
library(coefplot)

```

## Mean Salaries

The following text contains a descriptive analysis of the Salaries dataset, which contains the 6 variables Rank, Discipline, Years Since PhD, Years Since Service, Sex and Salary as well as 397 observations. Rank, Discipline and sex are categorial variables, Years Since PhD, Years Since Service and Salary are Integers. Remarkable about the Salary dataset is, that 90.18 percentage is male, that the salary is between 57,800 Euros and 231,545 Euros and that the Years Since Service differ between 0 and 60 years. 

Calculating the mean salaries by rank and sex shows that male university employees with the rank _Professor_ have the highest mean with 127,120.82 Euros, followed by the male _Association Professor_ with 94,869.70 Euros and the male _Assistant Professor_ with 81,311.46 Euros. Male employees mean is for all ranks higher than female employees mean, the salary for male employees is always above the mean and female's is always below the mean salary for each rank.

```{r, echo=FALSE} 
mean.salaries <- aggregate(salary~sex+rank, salaries, mean)
kable(mean.salaries, digits = 3, caption =  "Mean Salaries by sex and rank")
```

## Including Plots

Figure 1 shows a detailed description of the salary distribution by rank and sex in a Box-and-Whisker Plot. The turquoise boxplots represents the distribution of the male salaries by rank and the red boxplots the female salary distribution. Great differences are between female and male professor's and Assistant Professor's salary. 

```{r salary.boxplot, echo=FALSE, fig.cap="Displaying the salary distribution by sex and rank"}
boxplot(formula = salary ~ sex + rank,
        data = salaries,
        outline = TRUE,
        horizontal = TRUE,
        xlab = "Salaries in Euros",
        ylab = "Sex and rank of the university employees",
        col = rainbow(2))
```


According to the proportions, there are slightly more women in discipline _B_ than in discipline _A_. However, the same applies to men. This could mean that the number of employees in total in discipline _A_ is simply lower than the number of employees in discipline _B_. Because of that, it is not possible to draw a conclusion on whether the number of women differs based on discipline. 

```{r, echo=FALSE, warning=FALSE}
salary.discipline <- prop.table(table(salaries$sex, salaries$discipline))
kable(ftable(salary.discipline), digits = 3, caption = "proportions of women by discipline")

```

## Analysis of the Salary data with ANOVA

The mean salary for men and women among university employees is different. The one-way ANOVA displayed in table 3 shows that the F value is 7.738 and the p-value is very low --- less than 0.05. Hence, the conclusion is that the alternative hypothesis is accepted, the means of salaries for men and women are not equal, they are different.

```{r anova, echo=FALSE}
seg.aov.salary <- aov(salary ~ sex, data = salaries) 
aov.salary <- anova(seg.aov.salary)
kable(aov.salary, digits = 3, caption = "the results of the one-way ANOVA test")
```

The turkey post hoc test in figure 2 shows that the difference between salaries for men and women is statistically significant, since the p-value is smaller than 0.05. 

```{r anovaGraph, echo=FALSE, fig.cap="Testing and visualizing the ANOVA results on statistical signicficance"}
tuk <- TukeyHSD(seg.aov.salary)
plot(tuk)
```

## Linear model of the salary dataset

In order to bring salary closer to a normal distribution, salary was transformed using a log-function. Figure 3 shows the distribution of salary before transformation.

```{r transformationSalary, echo=FALSE, fig.cap="Plotting the salary of all observations"}
# theory http://fmwww.bc.edu/repec/bocode/t/transint.html
plot(salaries$salary)
transLog.salary <- log10(salaries$salary)
#transqrt.salary <- sqrt(salaries$salary)
```


The observations of the salary of the dataset shows a histogram in figure 4. The shape of the data is optimized for linear regression with a data transformation using the log-function. 

```{r, echo=FALSE, fig.cap="Transformed Salary with the log-function"}
# theory http://fmwww.bc.edu/repec/bocode/t/transint.html
salaries.df$salary <- log(salaries$salary)
hist(salaries.df$salary, xlab="Salary in Euro")
```

As shown in table 4, the standard error of the rank _AsstProf_ is the highest, which means that salaries for assistant professors differ by $167.315. The t-value indicates, whether a relationship between salary and e.g. sex or rank exists, whereas the value 0 means there is no relationship between the variables. There seems to be a clear relationship between salary and the rank _Prof_, as the t value is 9.290. With the value of _years since phd_ being so close to 0, there does not seem to be a relationship between _years since phd_ and salary.

```{r linearModel, echo=FALSE}
fit <- lm(formula = salary ~ sex + rank + discipline + yrs.since.phd,
          data = salaries) 
model <- data.frame(xtable(summary(fit)))
kable(model, caption = "summary of the linear model", digits = 3)

```

Figure 5 shows the coefficients for each variable. The figure shows once again, that there is no relationship between salary and male employees. However, the relationship between salary and employees with rank _Professor_ is evident.

```{r coefficientPlot, fig.cap="Visualization of the coefficinets of the linear model", echo=FALSE}
coefplot(fit, intercept = FALSE,
         outerCI = 1.96,
         lwdOuter = 1.5)
```


