---
title: "Special Work Performance 3: Diving into the Data"
author: 'Natalie und Julia'
geometry: margin=2cm
output: 
  pdf_document:
    fig_caption: true
    keep_tex: yes
sansfont: times
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(repos=c(CRAN = "https://cran.uni-muenster.de/"))
```

\begin{center}
Humboldt University Berlin 

course name und semester
by Prof. Dr. xxxxxxx 

\textbf{Group x:}\\  
\textbf{XXXXX}, 12345; 
\textbf{XXXXX}, 84579; \\
Submitted on November xx, xxxx 

\end{center}

```{r dataset, echo=FALSE}
## set environment
wdir <- "C:\\Users\\Natalie\\Dropbox\\WiSe201718\\Customer Analytics and Customer Insights\\Data"
setwd(wdir)
install.packages("psych"); library(psych)

## functions
std_rank <- function(x){
  return(sum(x)/(50*7))
}

## Reading data, once table "Attribute Rating", twice "Direct Preference Rating"
bars <- read.csv("attribute_evaluation_choc_bars.csv", sep = ";", header = TRUE)
rating <- read.csv("bars_rating.csv", sep = ";", header = TRUE)

rank <- list()
rating$Person <- NULL
rank <- sort(round(sapply(rating, std_rank), digits = 2), decreasing = TRUE)

## Cleaning sample bars
bars_mean <- list()
bars_mean <- sapply(na.omit(bars), mean)

for(col in 1:length(bars)) {
  #bars[[paste(col, "missing", sep = "_")]] <- ifelse(is.na(bars[[col]]), 1, 0)
  bars[[col]] <- ifelse(is.na(bars[[col]]), bars_mean[col], bars[[col]])
}
```

## Introduction of the Data Set "Chocolate Bars"

The given data set is about 10 different chocolate bar products. Four tables organize the accquired data in the tables "General Consumption", "Direct Preference Rating", "Attribute Rating" and "Social Demographic Questions". 15 participants evaluate 10 different choclate bars by 13 attributes. The participants rate directly their preferences of the chocolate bars on a scale from 1 till 7. The most popular bar is "Kinderriegel", the least popular one "Mars" as shown in figure 1.

```{r rank, echo=FALSE}
# main : rankings of chocolat bars
rank <- sort(round(sapply(rating, std_rank), digits = 2), decreasing = TRUE)
```

## Describing the Attributes of the Chocolate Bars

# Factor Analysis

The data reduction technique Factor Analysis investigates the relationships of the attributes by collapsing a large number of variables into a few interpretable underlying factors, the latent variables. The participants weight multiple variables similar because they belong to the same latent variable. 
In the first step, the data set "Attribute Rating", which contains the evaluation of each attribute per chocolat bar on a scale from 1 till 5 from 50 participants, is loaded and cleaned. The attribute's mean over all 50 observations replaces missing values. 

```{r fa, echo=FALSE}
## fa for each of the 10 chocolat products 
# here for the first chocolat product
b1_attr <- bars[,2:14]
cor(b1_attr)
#corrplot(b1_attr, method = "pie", type = "upper")
plot(eigen(cor(b1_attr))$values, type = "b", 
     main = "Plotting variances of Eigenvalues",
     xlab = "Attributes of Snickers",
     ylab = "Variance")   # use 6 factors

b1.fa <- fa(b1_attr, nfactors = 3, rotate = "variance", cut = 0.3)
fa.diagram(b1.fa)
#data.frame(b1.fa$scores)  #to-do: name columns
```

# Report and Interpret Results

The factor loading score shows now clear factorizing for the chocolate bar Snickers.
