---
title: "Special Work Performance 3: Reducing Data Complexity"
author: 'Natalie und Julia'
geometry: margin=2cm
output:
   pdf_document:
     latex_engine: xelatex
   fig_caption: true
   keep_tex: yes
#sansfont: times
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(repos=c(CRAN = "https://cran.uni-muenster.de/"))
```

\begin{center}
Humboldt University Berlin 

course name und semester
by Prof. Dr. xxxxxxx 

\textbf{Group x:}\\  
\textbf{XXXXX}, 12345; 
\textbf{XXXXX}, 84579; \\
Submitted on November xx, xxxx 

\end{center}

```{r dataset, echo=FALSE}
## set environment
#wdir <- "C:\\Users\\Natalie\\Dropbox\\WiSe201718\\Customer Analytics and Customer Insights\\Data"
#setwd(wdir)
#install.packages("psych"); 
library(psych)

## functions
std_rank <- function(x){
  return(sum(x)/(50*7))
}

## Reading data, once table "Attribute Rating", twice "Direct Preference Rating"
bars <- read.csv("attribute_evaluation_choc_bars.csv", sep = ";", header = TRUE)
rating <- read.csv("bars_rating.csv", sep = ";", header = TRUE)

rank <- list()
rating$Person <- NULL
rank <- sort(round(sapply(rating, std_rank), digits = 2), decreasing = TRUE)

## Cleaning sample bars
bars_mean <- list()
bars_mean <- sapply(na.omit(bars), mean)

for(col in 1:length(bars)) {
  #bars[[paste(col, "missing", sep = "_")]] <- ifelse(is.na(bars[[col]]), 1, 0)
  bars[[col]] <- ifelse(is.na(bars[[col]]), bars_mean[col], bars[[col]])
}
```

## Introduction of the Data Set "Chocolate Bars"

The given data set is about 10 different chocolate bar products. Four tables organize the accquired data in the tables "General Consumption", "Direct Preference Rating", "Attribute Rating" and "Social Demographic Questions". 50 participants evaluate the 10 different choclate bars Balisto Korn-Mix, Bounty, Duplo, Kinder Bueno, Kinder Riegel, KitKat, Lion, Mars by the 13 attributes sweet, handy, addiction, accessible, chocolaty, creamy, rich, wrapping, commercial, image, crunchy, healthful and calorie. The participants rate directly their preferences of the chocolate bars on a scale from 1 till 7. The most popular bar is _Kinderriegel_, the least popular one _Mars_ as shown in figure 1.

```{r rank, echo=FALSE, fig.cap="Figure 1: Rank of the 10 chocolat products"}
# main : rankings of chocolat bars
rank <- sort(round(sapply(rating, std_rank), digits = 2), decreasing = TRUE)
rank
```

## Describing the Attributes of the Chocolate Bars

### Euclidean Distances

The following presents the euclidean distances between the chocolate bars. In order to process the data accordingly, it was necessary to clean and transform the data. The objective is to create a distance matrix, which can then be used to visualise the distances on a two dimensional perceptual map. Using regression, the attributes should be mapped onto the perceptual map. The given version of the table _Attribute Rating_ contains one column for each chocolate bar - attribute - combination as well as one column _Person_. Hence, each row contains all the answers of one person. In order to produce the desired output, the data needed to be transformed so that there is an additional column _Chocolate Bar_ and only one column per attribute, e.g. crunchy, healthful, chocolaty etc. Furthermore, missing data needed to be removed by omitting rows that contain NA.

```{r euclidean distances, echo = FALSE}
library(knitr)
data <- na.omit(bars)

person <- c(data$Person, data$Person, data$Person, data$Person, data$Person, data$Person, data$Person, data$Person, data$Person, data$Person)
choc_bar <- rep(c("Snickers", "Kinder Bueno", "Twix", "Mars", "KitKat", "Bounty", "Kinder Riegel", "Balisto Korn-Mix", "Lion", "Duplo"), 
                each = length(data$Person))
crunchy <- c(data$b1_crunchy, data$b2_crunchy, data$b3_crunchy, data$b4_crunchy, data$b5_crunchy, data$b6_crunchy, data$b7_crunchy, data$b8_crunchy, data$b9_crunchy, data$bb_crunchy)
creamy <- c(data$b1_creamy, data$b2_creamy, data$b3_creamy, data$b4_creamy, data$b5_creamy, data$b6_creamy, data$b7_creamy, data$b8_creamy, data$b9_creamy, data$bb_creamy)
sweet <- c(data$b1_sweet, data$b2_sweet, data$b3_sweet, data$b4_sweet, data$b5_sweet, data$b6_sweet, data$b7_sweet, data$b8_sweet, data$b9_sweet, data$bb_sweet)
chocolaty <- c(data$b1_chocolaty, data$b2_chocolaty, data$b3_chocolaty, data$b4_chocolaty, data$b5_chocolaty, data$b6_chocolaty, data$b7_chocolaty, data$b8_chocolaty, data$b9_chocolaty, data$bb_chocolaty)
healthful <- c(data$b1_healthful, data$b2_healthful, data$b3_healthful, data$b4_healthful, data$b5_healthful, data$b6_healthful, data$b7_healthful, data$b8_healthful, data$b9_healthful, data$bb_healthful)
calorie <- c(data$b1_calorie, data$b2_calorie, data$b3_calorie, data$b4_calorie, data$b5_calorie, data$b6_calorie, data$b7_calorie, data$b8_calorie, data$b9_calorie, data$bb_calorie)
rich <- c(data$b1_rich, data$b2_rich, data$b3_rich, data$b4_rich, data$b5_rich, data$b6_rich, data$b7_rich, data$b8_rich, data$b9_rich, data$bb_rich)
addiction <- c(data$b1_addiction, data$b2_addiction, data$b3_addiction, data$b4_addiction, data$b5_addiction, data$b6_addiction, data$b7_addiction, data$b8_addiction, data$b9_addiction, data$bb_addiction)
accessible <- c(data$b1_accessible, data$b2_accessible, data$b3_accessible, data$b4_accessible, data$b5_accessible, data$b6_accessible, data$b7_accessible, data$b8_accessible, data$b9_accessible, data$bb_accessible)
handy <- c(data$b1_handy, data$b2_handy, data$b3_handy, data$b4_handy, data$b5_handy, data$b6_handy, data$b7_handy, data$b8_handy, data$b9_handy, data$bb_handy)
wrapping <- c(data$b1_wrapping, data$b2_wrapping, data$b3_wrapping, data$b4_wrapping, data$b5_wrapping, data$b6_wrapping, data$b7_wrapping, data$b8_wrapping, data$b9_wrapping, data$bb_wrapping)
image <- c(data$b1_image, data$b2_image, data$b3_image, data$b4_image, data$b5_image, data$b6_image, data$b7_image, data$b8_image, data$b9_image, data$bb_image)
commercial <- c(data$b1_commercial, data$b2_commercial, data$b3_commercial, data$b4_commercial, data$b5_commercial, data$b6_commercial, data$b7_commercial, data$b8_commercial, data$b9_commercial, data$bb_commercial)

data_cleaned <- data.frame(person, choc_bar, crunchy, creamy, sweet, chocolaty, healthful, calorie, rich, addiction, accessible, handy, wrapping, image, commercial)
#data_cleaned <- na.omit(data_complete)
kable(data[1:2, 1:4], caption = "data before transformation")

kable(data_cleaned[1:2, 1:5], caption = "data after transformation")
```

Then, the data was aggregated by calulating the mean of each combination of chocolate bars and attributes.

```{r, echo=FALSE}
data_aggregated <- aggregate(data_cleaned[,-c(1,2)], by=list(data_cleaned$choc_bar),mean, na.rm=TRUE)
kable(data_aggregated[1:2, 1:4], caption = "mean values of chocolate bar - attribute - combinations")
```

The resulting table was used to create a matrix containing euclidean distances between the chocolate bars. In order to visualize the data on a two dimensional plane, multidimensional scaling was necessary. Using regression, the attributes were then fitted into the map. For visibility purposes, the upper table gives a general view on the distribution of the data on a two dimensional plane, whereas the lower table concentrates on the center of the map, where most of the attributes are located.  

```{r, echo=FALSE}
distance_df <- as.matrix(dist(data_aggregated[,-c(1)], method = 'euclidean'), labels=TRUE)
colnames(distance_df) <- rownames(distance_df) <- data_aggregated$Group.1
fit <- cmdscale(distance_df)
x <- fit[,1]
y <- fit[,2]

data_aggregated$x <- x
data_aggregated$y <- y

profit <- lm(cbind(crunchy, creamy, sweet, chocolaty, healthful, calorie,
                   rich, addiction, accessible, handy, wrapping, image, commercial) ~ -1 + x + y, data = data_aggregated)

plot(x, y, xlab = "Coordinate 1", ylab = "Coordinate 2", main = "Metric MDS for Chocolate Bars", pch = 19, ylim = c(-3, 3), xlim = c(-3, 3))
text(x, y, labels = data_aggregated$Group.1, cex = 0.75, pos = 4) 
abline(h = 0, v = 0, col = "grey")
arrows(x0 = c(0, 0, 0), y0 = c(0, 0, 0),
       x1 = coef(profit)[1, ]*3, y1 = coef(profit)[2, ]*3, col = 2, lwd = 1)
text(t(coef(profit)*3.5), colnames(coef(profit)*3.5), cex=0.6, col = 1, pos = 4)

```

```{r, echo=FALSE}
plot(x, y, xlab = "Coordinate 1", ylab = "Coordinate 2", main = "Metric MDS for Chocolate Bars 2", pch = 19, ylim = c(-1.5, 1.5), xlim = c(-1.5, 1.5))
text(x, y, labels = data_aggregated$Group.1, cex = 0.75, pos = 4) 
abline(h = 0, v = 0, col = "grey")
arrows(x0 = c(0, 0, 0), y0 = c(0, 0, 0),
       x1 = coef(profit)[1, ]*3, y1 = coef(profit)[2, ]*3, col = 2, lwd = 1)
text(t(coef(profit)*3), colnames(coef(profit)*3), cex=0.6, col = 1, pos = 4)
```

The tables show that the chocolate bars Snickers, Twix, Duplo and KitKat are rated more crunchy, whereas Kinder Riegel, Mars and Bounty are rated creamy. It is also visible that none of the chocolate bars have been rated very healthful - the highest rated is Balisto Korn-Mix with mean = 2.17. Altogether, the chocolate bars are distributed quite evenly on the map, which shows that the differences according to the attributes are rather small. For each of the attributes in the center of the map (handy, addiction, sweet, calorie, wrapping, rich), the chocolate bars have very similar ratings. The ratings for attributes such as creamy and crunchy, however, are more diverse.

### Factor Analysis

The data reduction technique Factor Analysis investigates the relationships of the attributes by collapsing a large number of variables into a few interpretable underlying factors, the latent variables. The data reduction technique Principal Component Analysis summarizes attributes to subgroups. Becuase principal component analysis simply reduce the amount of given variables, factor analysis is applied here to discover latent variable for a better understanding which group of variables are important for the consumer in general for a chocolate bar. 
The participants weight multiple variables similar because they belong to the same latent variable. The question that should be answered with factor analysis in the following analysis is "What are the important attributes in general of a chocolat product?". The findings show which factors have importance for the consumer. 
Firstly, the data set is cleaned and analyzed before using the factor analysis function. The data set "Attribute Rating", which contains the evaluation of each attribute per chocolat bar on a scale from 1 till 5 from 50 participants, is loaded and cleaned. The cleaning program calculates the attribute's mean over all 10 chocolate products. The result is a 50 by 13 matrix.  

Factor analysis assign each variable to a latent variable, the factor, based on their correlation. The value of the eigenvalues of the data set as a matrix show how many variables the eigenvector explains. Fig. B1 shows the eigenvalue of each variable. 5 variables have a eigenvalue > 1. The model has a good fit with 5 or less factors.

```{r eigenvalue, echo=FALSE, fig.cap="Fig. B1: Sorted list of the eigenvalues of the Participants-Attribute Matrix"}
if(!require("tidyverse")) install.packages("tidyverse"); library("tidyverse") # load the package
if(!require("lavaan")) install.packages("lavaan",dependencies=T); library("lavaan") # load the package
if(!require("semTools")) install.packages("semTools", dependencies=T); library("semTools") # load the package
# if(!require("semPlot")) install.packages("semPlot", dependencies=T); library("semPlot") # load the package
if(!require("psych")) install.packages("psych", dependencies=T); library("psych") # load the package
if(!require("car")) install.packages("car", dependencies=T); library("car") # load the package
if(!require("RColorBrewer")) install.packages("RColorBrewer", dependencies=T); library("RColorBrewer") # load the package

attributeRating_df <- read.csv("attribute_evaluation_choc_bars.csv", sep = ";", header = TRUE)

attributeRating_df <- attributeRating_df %>% 
  gather(k, v, -Person) %>% 
  separate(k, c("brand","b"), sep =  "_") %>% 
  spread(b, v, fill = "") %>% 
  arrange(brand)

attributeRating_df[, 3:15] <- sapply(attributeRating_df[, 3:15], as.numeric)
attributeRating_df$brand <- factor(attributeRating_df$brand)
levels(attributeRating_df$brand) <- c("Snickers", "Kinder Bueno", "Twix", "Mars", "KitKat",
                                    "Bounty", "Kinder Riegel", "Balisto Korn-Mix", "Lion",
                                    "Duplo")

#summary(attributeRating_df[, 3:15])
#describe(attributeRating_df[, 3:15], na.rm = TRUE)

sort(round(eigen(cor(attributeRating_df[,3:15], use = "complete.obs"))$values, 2), decreasing = TRUE) #result: 5 factors EV > 1
```

### Report Results

The model best fit is with 5 factors, using the method "minres" and the orthogonal rotation model. One variable is assigned to one factor with the lower theshold at 0.3. The first factor induces the attributes image, rich, accessible and wrapping. The assigned name is _Packaging_. The attribute rich is assigned with 0.3 which explains the weak relationship to the factor _Packaging_. The second factor induces the attributes image and commercial. The assigned name is _Brand_. The third factor induces the attributes chocolaty, addiction and sweet, the name of the factor is  _Taste_. The fourth factor induces the variables creamy and crunchy. The assigned name is  _Consistency_. The fifth factor induces the attributes healthful and calorie. The assigned name is _Health Awareness_. Healthful is assigned with 0.6 and calorie with -0.5. It shows that the two attributes are antiproportional assigned to the factor. Fig. B2 shows the scores of the first two factors _Packaging_ and _Brand_. 

```{r fa, echo=FALSE, fig.cap="Fig. B2: Factor scores of the factors Packaging and Taste"}
pies.fa <- fa(attributeRating_df[3:15], missing = TRUE, nfactors = 5, rotate = "varimax", cut = 0.3)
#fa.diagram(pies.fa)
colnames(pies.fa$scores)[1] <- "Packaging"
colnames(pies.fa$scores)[2] <- "Brand"
fa.scores <- data.frame(pies.fa$scores)[1:10,1:2]
round(fa.scores, 2)
```

## Interpret results

In the euclidean distance calculation the attributes _commercial_ and _image_ build one of the most important group in order to differentiate the different chocolate bars. _Commercial_ and _image_ induce the factor _Brand_ in the factor analysis model. Because of the high correlation the attributes express only slightly different consumers preferences and are substituteable or summarized to one attribute.
The attributes _crunchy_ and _creamy_ differentiate the chocolat products in opposite groups. In the factor analysis model this differentitation is concluded in the factor _consistency_. However, the attribute _creamy_ has a higher correlation than _crunchy_. In the MDS model it is the other way around. The ratings for the attributes _wrapping_, _handy_, _accesible_, _rich_ and _image_ are quite similar, so it is reasonable to combine them to a factor _packaging_, as suggested by the factor analysis. Although fewer attributes contribute to clarity and comprehensibility, crucial attributes, such as taste-attributes for chocolate bars, should not be reduced. Taste is extremely subjective, hence it is more expressive to rate attributes such as _creamy_ or _chocolaty_. The metric MDS is quite dense around the center with a relatively even distribution of attributes. With less attributes, the metric MDS would be much more readable. Pre-processing the data by performing a factor analysis first would be a reasonable approach in this case.  

## Appendix
```{r, echo=FALSE}
library(pander)
pander(distance_df, split.table = 95, style = 'rmarkdown', caption = "distance matrix for chocolate bars")
pander(fit, split.table=95, style='rmarkdown', caption='coordinates after multidimensional scaling using cmdscale')
#kable(distance_df)
```

Factor Analysis: Diagram of the assignment of the variables to one factor

```{r appendix_fa, echo=FALSE, fig.cap="Diagram of Factor Analysis"}
fa.diagram(pies.fa)
```
