---
title: 'Special Work Performance 3: Diving into the Data'
author: "Natalie und Julia"
geometry: margin=2cm
header-includes:
  - \usepackage{tinytex}
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
  html_document:
    df_print: paged
sansfont: times
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(repos=c(CRAN = "https://cran.uni-muenster.de/"))
```
\begin{center}
Humboldt University Berlin 

course name und semester
by Prof. Dr. xxxxxxx 

\textbf{Group x:}\\  
\textbf{XXXXX}, 12345; 
\textbf{XXXXX}, 84579; \\
Submitted on November xx, xxxx 

\end{center}

```{r dataset, echo=FALSE}
## set environment
wdir <- "C:\\Users\\Natalie\\Dropbox\\WiSe201718\\Customer Analytics and Customer Insights\\Data"
setwd(wdir)
install.packages("psych"); 
library(psych)

## functions
std_rank <- function(x){
  return(sum(x)/(50*7))
}

## Reading data, once table "Attribute Rating", twice "Direct Preference Rating"
bars <- read.csv("attribute_evaluation_choc_bars.csv", sep = ";", header = TRUE)
rating <- read.csv("bars_rating.csv", sep = ";", header = TRUE)

rank <- list()
rating$Person <- NULL
rank <- sort(round(sapply(rating, std_rank), digits = 2), decreasing = TRUE)

## Cleaning sample bars
bars_mean <- list()
bars_mean <- sapply(na.omit(bars), mean)

for(col in 1:length(bars)) {
  #bars[[paste(col, "missing", sep = "_")]] <- ifelse(is.na(bars[[col]]), 1, 0)
  bars[[col]] <- ifelse(is.na(bars[[col]]), bars_mean[col], bars[[col]])
}
```

## Introduction of the Data Set "Chocolate Bars"

The given data set is about 10 different chocolate bar products. Four tables organize the accquired data in the tables "General Consumption", "Direct Preference Rating", "Attribute Rating" and "Social Demographic Questions". 15 participants evaluate 10 different choclate bars by 13 attributes. The participants rate directly their preferences of the chocolate bars on a scale from 1 till 7. The most popular bar is "Kinderriegel", the least popular one "Mars" as shown in figure 1.

```{r rank, echo=FALSE}
# main : rankings of chocolat bars
rank <- sort(round(sapply(rating, std_rank), digits = 2), decreasing = TRUE)
```

## Describing the Attributes of the Chocolate Bars

#Euclidean Distances

The following presents the euclidean distances between the chocolate bars. In order to process the data accordingly, it was necessary to clean and transform the data. The objective is to create a distance matrix, which can then be used to visualise the distances on a two dimensional perceptual map. Using regression, the attributes should be mapped onto the perceptual map. The given version of the table _Attribute Rating_ contains one column for each chocolate bar - attribute - combination as well as one column _Person_. Hence, each row contains all the answers of one person. In order to produce the desired output, the data needs to be transformed so that there is an additional column _Chocolate Bar_ and only one column per attribute, e.g. crunchy, healthful, chocolaty etc. \\
\begin{table}
person &    b1_crunchy &  b1_creamy  &     ...  &     b2_crunchy \\
\hline
 1     &    5         &   4       & ...         &      4
\end{table}

  person     b1_crunchy   b1_creamy       ...       b2_crunchy
 -------     ----------  ----------      ----       ----------
  1                   5           4       ...                4                     

  person     chocolate bar   crunchy    creamy       ...    b2_crunchy
 -------            ------  --------   -------   -------       -------
       1          Snickers         4         5       ...             4   


# Factor Analysis

The data reduction technique Factor Analysis investigates the relationships of the attributes by collapsing a large number of variables into a few interpretable underlying factors, the latent variables. The participants weight multiple variables similar because they belong to the same latent variable. 
In the first step, the data set "Attribute Rating", which contains the evaluation of each attribute per chocolat bar on a scale from 1 till 5 from 50 participants, is loaded and cleaned. The attribute's mean over all 50 observations replaces missing values. 
-Step 2: Examine correlation matrix
-Step 3: Determine number of factors by eigenvalue >1
-Step 4: Choose a factor rotation model, here rotation = "varimax"
-Step 5: fit the model and interpret the factors, assign them with names
-Step 6: Repeat and compare with different factor levels, rotation method
-Step 7: Use factor scores for respondents, factor scores show the best estimate for the latent variable

```{r fa, echo=FALSE}
## fa for each of the 10 chocolat products 
# here for all 10 chocolat products, one fa per chocolat product
ctr <- 2
while(ctr < length(bars)) {
  last_attr <-ctr+12
  bar_attr <- bars[,ctr:last_attr]
  #cor(bar_attr)
  nfac <- length(eigen(cor(b1_attr))$values[eigen(cor(b1_attr))$values > 1.0])
  b.fa <- fa(bar_attr, nfactors = nfac, rotate = "varimax", cut = 0.3)
  fa.diagram(b.fa)
  data.frame(b.fa$scores)[1:2,]
  
  ctr <- last_attr+1
}

```

# Report and Interpret Results


