---
title: "Special Work Performance 1"
geometry: margin=2cm
output: 
  pdf_document:
    fig_caption: true
sansfont: times
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

1. there are 6 variables and 397 observations in the data set
```{r}
library(car)
library(corrplot)
data <- Salaries
dimensions <- dim(data)
```
2. there are 19 professors who have more than 40 yeats of service, assuming the term \textit{professor} refers only to observations with rank == 'Prof'
```{r}
data.df <- data.frame(data)
num_of_profs <- sum(subset(data.df$rank == 'Prof', data.df$yrs.service > 40))
```
3. assuming the rank is not specified, 54 people have salary larger than $150,000. 
```{r}
salary <- nrow(subset(data, salary > 150000))
```
4. the mean salary for professors with more than 20 years of service is $126,772.1 with the rank restricted to rank == 'Prof'
```{r}
mean_salary <- mean((subset(data$salary, data.df$rank == 'Prof', data.df$yrs.service > 20)))
```
5. you can find out more about the data set by using the function summary():
```{r}
summary(data)
```
6. the following tables provide the counts and proportions of men and women by rank. The results show that the number of male professors is much higher than the number of women.  
```{r echo=TRUE}
tab <- table(data$sex, data$rank)
knitr::kable(
  head(tab[, 1:3], 3), caption = 'counts of men and women by rank'
)
tab2 <- prop.table(tab)
knitr::kable(
  head(tab2[, 1:3], 3), caption = 'proportions of men and women by rank'
)
```

7. the following figure provides a histogram for years of service.
```{r fig1, echo=TRUE, fig.cap="Histogram of years of service for both men and women"}
hist(data$yrs.service, freq=FALSE, xlab="years of service",
     ylab="proportional density", main="Histogram of years of service")
lines(density(data$yrs.service), type="l", col="darkred")
```


